Unshuffling Permutations

### 2015-09-16 article.tex
nouvelle version avec les preuves qui manquaient

For a graph $G$, we denote $\VS{G}$ as the set of vertices and
$\ES{G}$ as the set of edges.
Let $u = u_1 \, u_2 \,\ldots\, u_n$ be a word.
The \emph{clique graph associated to $u$}, denoted $K_u$,
is the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$.
The structure of this underlying graph is linear,
\emph{i.e.}, the set of vertices is equipped with a natural total order
$<$ defined by $u_i < u_j$ if and only if $i < j$.
(For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
for an edge if it is clear from the context that $i < j$.)
Recall that two edges of a graph are \emph{independent} if they do not share
a common vertex, and that a \emph{matching} $\mathcal{M}$ in $G$ is a set of
pairwise independent edges.
A matching is \emph{perfect} if it covers all the vertices of the graph.
In case the set of vertices is equipped with a total order,
a matching $\mathcal{M}$ is said to be \emph{containment-free}
if there do not exist (independent) edges $(u_i, u_j)$ and
$(u_k, u_\ell)$ in $\mathcal{M}$ such that
$u_i < u_k < u_\ell < u_j$ or  $u_k < u_i < u_j < u_\ell$.

@@

  (\emph{2} $\Rightarrow$ \emph{1}).
  Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
  \;\text{with}\; j < k\}$, and
  $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
  \;\text{with}\; j < k\}$.
  Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right) endpoints
  of the edges in $\mathcal{M}$ assuming the positions $1, 2, \ldots, 2$
  are drawn along an horizontal line.
  Without loss of generality,
  assume $i^1_1 < i^1_2 < \ldots < i^1_n$.
  % and
  % $I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$.
  Let
  $p^1 = \pi(i^1_1) \pi(i^1_2) \ldots \pi(i^1_n)$ and
  $p^2 = \pi(i^2_1) \pi(i^2_2) \ldots \pi(i^2_n)$ be the two patterns
  from $\pi$ induced by $I^1$ and $I^2$, respectivelly.
  Clearly $p^1$ and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is
  a matching) and cover $\pi$ (since $\mathcal{M}$ is perfect).
  We also claim that $p^1$ and $p^2$ are order-isomorphic.
  Suppose, aiming at a contradiction, that this is not the case.
  Therefore, there exist two positions $j$ and $k$, say $j < k$, such that
  $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
  $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
  But since $\mathcal{M}$ is containment-free, it follows that
  both $\{i^1_j, i^2_j\}$ and $\{i^1_k, i^2_k\}$.
  This is a contradiction since
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\pi(a) < \pi(a')$ if and only if $\pi(b) < \pi(b')$.
  \qed
\end{proof}

>>>>

For a graph $G$, we denote $\VS{G}$ as the set of vertices and
$\ES{G}$ as the set of edges.
Let $u = u_1 \, u_2 \,\ldots\, u_n$ be a word.
The \emph{clique graph associated to $u$}, denoted $K_u$,
is the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$.
The structure of this underlying graph is linear,
\emph{i.e.}, the set of vertices is equipped with a natural total order
$<$ defined by $u_i < u_j$ if and only if $i < j$.
For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
for an edge if it is clear from the context that $i < j$.
Recall that two edges of a graph are \emph{independent} if they do not share
a common vertex, and that a \emph{matching} $\mathcal{M}$ in $G$ is a set of
pairwise independent edges.
A matching is \emph{perfect} if it covers all the vertices of the graph.
In case the set of vertices is equipped with a total order,
a matching $\mathcal{M}$ is said to be \emph{containment-free}
if there do not exist (independent) edges $(u_i, u_j)$ and
$(u_k, u_\ell)$ in $\mathcal{M}$ such that
$u_i < u_k < u_\ell < u_j$ or  $u_k < u_i < u_j < u_\ell$.

@@

  (\emph{2} $\Rightarrow$ \emph{1}).
  Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
  \;\text{with}\; j < k\}$, and
  $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
  \;\text{with}\; j < k\}$.
  Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right) endpoints
  of the edges in $\mathcal{M}$ assuming the positions $1, 2, \ldots, 2$
  are drawn along an horizontal line.
  Without loss of generality,
  assume $i^1_1 < i^1_2 < \ldots < i^1_n$.
  % and
  % $I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$.
  Let
  $p^1 = \pi(i^1_1) \pi(i^1_2) \ldots \pi(i^1_n)$ and
  $p^2 = \pi(i^2_1) \pi(i^2_2) \ldots \pi(i^2_n)$ be the two patterns
  from $\pi$ induced by $I^1$ and $I^2$, respectivelly.
  Clearly $p^1$ and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is
  a matching) and cover $\pi$ (since $\mathcal{M}$ is perfect).
  We also claim that $p^1$ and $p^2$ are order-isomorphic.
  Suppose, aiming at a contradiction, that this is not the case.
  Therefore, there exist two positions $j$ and $k$, say $j < k$, such that
  $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
  $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
  But since $\mathcal{M}$ is containment-free, it follows that
  both $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and
  $\{i^1_k, i^2_k\}$, $i^1_k, i^2_k$, belong to $\mathcal{M}$.
  This is a contradiction since
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\pi(a) < \pi(a')$ if and only if $\pi(b) < \pi(b')$.
  \qed
\end{proof}
